```{r setup}
library(ggplot2)
library(tidyverse)
library(broom)
library(reticulate)
library(pracma)
library(cowplot)
source("scripts/curve_fitting.R")
# py_install("pandas")
source_python("read_pickle.py")

```


```{r}
# Download a file for analysis
df_sensor_sn <- read_pickle_file(
    "./hepa-pckls/indoor-outdoor-pairs/48 Putnam Street #1/MOD-PM-00112.pckl"
)
```


```{r}
# Convert all zeros in outdoor data to NA to prevent divide by zero errors
bin <- na.omit(str_match(names(df_sensor_sn), "^outdoor_([neph|opc].*)")[,1])
df_sensor_sn[, bin][df_sensor_sn[, bin] < 0.01] <- NA

# Add columns for sub 1nm particle count
df_sensor_sn["outdoor_particle_sum"] <- df_sensor_sn["outdoor_opc_bin0"] + df_sensor_sn["outdoor_opc_bin1"] + df_sensor_sn["outdoor_opc_bin2"]
df_sensor_sn["indoor_particle_sum"] <- df_sensor_sn["indoor_opc_bin0"] + df_sensor_sn["indoor_opc_bin1"] + df_sensor_sn["indoor_opc_bin2"]

```


```{r}
# Add Indoor / Outdoor ratio columns to the dataframe

# Create a list of data types
bin <- na.omit(str_match(names(df_sensor_sn), "^indoor_(.*)")[,2])

# Create list of cols from original data 
indoor <- df_sensor_sn[paste0("indoor_", bin)]
outdoor <- df_sensor_sn[paste0("outdoor_", bin)]

#create new col names
ratio_name <- paste0("ratio_", bin)

# Calculate new cols and add to dataframe
df_sensor_sn[ratio_name] <- indoor / outdoor

df_sensor_sn %>%
  unnest_wider(hepa_installed)

```

```{r}
df_sensor_sn %>%
  group_by(hepa_installed) %>%
  ggplot(aes(timestamp, indoor_opc_bin0)) +
  geom_line(color = "blue") +
  geom_line(aes(y = outdoor_opc_bin0), color = "green") + 
  geom_line(aes(y = ratio_opc_bin0)) + 
  geom_hline(yintercept = quantile(df_sensor_sn$indoor_opc_bin0, .95, na.rm = TRUE ), linetype="dashed", color = "red") +
  facet_wrap(vars(hepa_installed))
```

Exposure: Concentration / time (area under curve)

Concentration: Amount fo pollutant in the air

modpm opc bins are in particles / m^3 not ug/m^3

Indoor/Outdoor ratio: how do we account for this with above metrics?

```{r}
# Find statistics for before and after
df_hepa_stats <-
  df_sensor_sn %>%  
  group_by(hepa_installed) %>%
  summarize(across(contains("bin"),
                   list(fifth = ~ quantile(.x, .05, na.rm = TRUE),
                        twentyfifth = ~ quantile(.x, .25, na.rm = TRUE),
                        seventyfifth = ~ quantile(.x, .75, na.rm = TRUE),
                        nintyfifth = ~ quantile(.x, .95, na.rm = TRUE),
                        mean = ~ mean(.x, na.rm = TRUE),
                        median = ~ median(.x, na.rm = TRUE)
                        )
                   )
            ) %>%
  pivot_longer(
    names_to = c("location", "sensor", "bin", ".value"),
    names_sep = "_",
    cols = -c(hepa_installed)
  ) %>%
  filter(location != "outdoor")

df_hepa_stats

```


```{r}
df_sensor_sn %>%
  filter(hepa_installed == FALSE) %>%
  ggplot(aes(timestamp, indoor_opc_bin0)) +
  geom_line(color = "blue") +
  geom_hline(yintercept = df_hepa_stats$fifth[1], linetype = "dashed", color = "red") + 
  geom_hline(yintercept = df_hepa_stats$twentyfifth[1], linetype = "dashed", color = "green") +
  geom_hline(yintercept = df_hepa_stats$median[1], linetype = "dashed", color = "black") + 
  geom_hline(yintercept = df_hepa_stats$seventyfifth[1], linetype = "dashed", color = "orange")

df_sensor_sn %>%
  filter(hepa_installed == TRUE) %>%
  ggplot(aes(timestamp, indoor_opc_bin0)) +
  geom_line(color = "blue") +
  geom_hline(yintercept = df_hepa_stats$fifth[2], linetype = "dashed", color = "red") + 
  geom_hline(yintercept = df_hepa_stats$twentyfifth[2], linetype = "dashed", color = "green") +
  geom_hline(yintercept = df_hepa_stats$median[1], linetype = "dashed", color = "black") + 
  geom_hline(yintercept = df_hepa_stats$seventyfifth[1], linetype = "dashed", color = "orange")
```
25th percentile seems like best metric for baseline concentration, therefore the
reduction of baseline concentration can be found by comparing the 25th percentile
before and after installation. This is a time independent metric but it requires
us to ignore potential impacts of outdoor concentrations.

indoor/outdoor ratio baseline reduction is less compelling and harder to understand

```{r}
df_baseline_reduction <-
  df_hepa_stats %>%
  group_by(location, sensor, bin) %>%
  summarise(baseline_reduction = (median[1] - median[2]))

df_baseline_reduction
```
```{r}
# 
df_baseline_reduction["binint"] <- as.integer(sub(".*?(\\d+)", "\\1", df_baseline_reduction$bin))

df_baseline_reduction %>%
  filter(location == "indoor", sensor == "opc") %>%
  ggplot(aes(binint, baseline_reduction, label = baseline_reduction)) +
  geom_point()+
  geom_text(hjust = 0, vjust = -1)
  xlab("Bin Number")



```

```{r}

```

```{r}

```






