---
title: "HAFTRAP_R_Notebook"
output: html_notebook
author: Scott Hersey, drawing from Neel Dhulipala's HEPA-data-analysis code in places.
---



```{r}
library(openair)
library(corrplot)
library(ggplot2)
library(reshape2)
library(purrr)
library(tidyr)
library(lubridate)
library(reticulate)
library(dplyr)

# Define files to be imported and filter on datetime
indoorfile <- 'data/MOD_indoor.csv'
outdoorfile <- 'data/MOD_outdoor.csv'

hepatime <- "2022-02-07 10:00:00"
  
# HEPA Times
# Deployment 1: datetime(2022,02,07,10,00,00)
# Deployment 2: datetime(2022,02,16,10,00,00)
# Deployment 3: datetime(2022,03,20,08,00,00)
# Deployment 4: datetime(2022,04,08,09,00,00)
```


```{r}
# Import and prepare data

# Import indoor data
indoordata <- read.csv(indoorfile, header = TRUE)
date <- as.POSIXct(strptime(indoordata$timestamp_local, format = "%Y-%m-%dT%H:%M:%SZ", tz = "America/New_York"))
indoordata$date_1min <- round_date(date, unit = "1 min") #creating date1min vector. Turns 00:00:01 and 00:00:59 into 00:00 and 00:01, respectively.
indoordata$pm1_num <- indoordata$bin0 + indoordata$bin1 + indoordata$bin2
#indoordata <- cbind(indoordata, date, date_1min, pm1_num)
#indoordata <- cbind(indoordata, date) #add back in a "date" column for openair functions

# Import outdoor data
outdoordata <- read.csv(outdoorfile, header = TRUE)
date <- as.POSIXct(strptime(outdoordata$timestamp_local, format = "%Y-%m-%dT%H:%M:%SZ", tz = "America/New_York"))
outdoordata$date_1min <- round_date(date, unit = "1 min") #creating date1min vector. Turns 00:00:01 and 00:00:59 into 00:00 and 00:01, respectively.
outdoordata$pm1_num <- outdoordata$bin0 + outdoordata$bin1 + outdoordata$bin2
#outdoordata <- cbind(outdoordata, date, date_1min, pm1_num)

#outdoordata <- cbind(outdoordata, date) #add back in a "date" column for openair functions

# Set 0 values in outdoor data to NaN to prevent errors in creating ratios later on
outdoordata[outdoordata == 0] <- NA
```


```{r}
# Eliminate columns that indoor and outdoor dataframes don't have in common

# extract column names of the data frames
cols_indoor <- names(indoordata)
cols_outdoor <- names(outdoordata)

# find the intersection and difference of both column name vectors
cols_intersection <- intersect(cols_indoor, cols_outdoor)
cols_diff <- setdiff(cols_indoor, cols_outdoor)

# subset the initial data frames
indoordata <- indoordata[,cols_intersection]
outdoordata <- outdoordata[,cols_intersection]
```


```{r}
# Label all variables by "outdoor_" or "indoor_"
colnames(outdoordata) <- paste0('outdoor_', colnames(outdoordata)) # add "outdoor_" label to all headers
colnames(indoordata) <- paste0('indoor_', colnames(indoordata)) # add "indoor_" label to all headers


# Time sync indoor and outdoor data
joineddata <- left_join(indoordata, outdoordata, by = c("indoor_date_1min" = "outdoor_date_1min"))
joineddata$date <- joineddata$indoor_date_1min #add back in a "date" column for openair functions

# Attach joineddata to make calling variables easier
attach(joineddata)
```



```{r}
#Initial time series
class(date)
timePlot(joineddata, pollutant = c("indoor_bin0", 'indoor_pm1', 'indoor_pm25'))
timePlot(joineddata, pollutant = c("outdoor_bin0", 'outdoor_pm1', 'outdoor_pm25'))

class(date)
timePlot(joineddata, pollutant = "indoor_bin0")

```


```{r}
# Create some initial diurnal plots. To be replaced by static report code for mean/median/percentile plots.
#date <- indoordata$indoor_date
timeVariation(joineddata, pollutant = c("indoor_no", "indoor_co")) 
timeVariation(joineddata, pollutant = c("indoor_pm1", "indoor_pm25"))

timeVariation(joineddata, pollutant = c("outdoor_no", "outdoor_co")) 
timeVariation(joineddata, pollutant = c("outdoor_pm1", "outdoor_pm25"))
```


```{r}
# Correlation plots for variables
pdf(file = "figures/corplot.pdf",   # The directory you want to save the file in
    width = 8, # The width of the plot in inches
    height = 10) # The height of the plot in inches


cordata <- data.frame( indoor_temp,  indoor_rh,  indoor_pm1_num,  indoor_opcn3_pm1,  indoor_opcn3_pm25,  indoor_opcn3_pm10,  indoor_pm1_env,  indoor_pm25_env,  indoor_pm10_env,  indoor_neph_bin0,  indoor_no,  indoor_co,  indoor_no2,  indoor_pm25,  indoor_pm10, outdoor_temp,  outdoor_rh,  outdoor_pm1_num,  outdoor_opcn3_pm1,  outdoor_opcn3_pm25,  outdoor_opcn3_pm10,  outdoor_pm1_env,  outdoor_pm25_env,  outdoor_pm10_env,  outdoor_neph_bin0,  outdoor_no,  outdoor_co,  outdoor_no2,  outdoor_pm25,  outdoor_pm10) # indoor_o3

#outdoorcordata <- data.frame( outdoor_temp,  outdoor_rh,  outdoor_pm1_num,  outdoor_opcn3_pm1,  outdoor_opcn3_pm25,  outdoor_opcn3_pm10,  outdoor_pm1_env,  outdoor_pm25_env,  outdoor_pm10_env,  outdoor_neph_bin0,  outdoor_no,  outdoor_co,  outdoor_no2,  outdoor_pm25,  outdoor_pm10) #joineddata$outdoor_o3

#cordata <- cbind(indoorcordata, outdoorcordata)

N <- cor(na.omit(cordata))
corrplot(N, method = 'pie')

```

```{r}
# importing matplotlib for diurnal plots

# IF you get a ModuleNotFoundError with any packages from diurnal_plot.py,
# replace "matplotlib" with that package (e.g. pandas, numpy, etc.)
py_install(packages = "matplotlib")
```

```{r}
# Create timestamp variable for diurnal plots
joineddata$timestamp <- joineddata$indoor_date_1min
```

```{python}
# Create diurnal plot with summary statistics for dataframe of air quality
# and particulate matter
import pandas as pd
import diurnal_plot as dp

df = r.joineddata
plot = dp.DiurnalPlot("outdoor_pm1")
dataPM = plot.process_data(df)
dataPM = dataPM.groupby('time')["outdoor_pm1"]
dataPM.mean()

# dp.diurnal_plot(r.joineddata, "outdoor_pm1")
```

```{r}
# Calculate indoor/outdoor ratios
# NOTE: This works, and cycles through all of the numeric variables. Somehow it gets stuck somewhere at the very end, but it does work. Potential solution is to create new indoor/outdoor dfs based on the indoor/outdoor labels and compare the headers to see if something is there in indoor but not there in outdoor?



# Go through every column of the dataframe
for (colname in names(indoordata)) {
    if (grepl("indoor_", colname, fixed = TRUE)) {
        datatype <- substring(colname, 8, nchar(colname))  # Determines what the data vector is without the "indoor_" label
        # Get indoor and outdoor columns that includes same datatype
        indoor <- joineddata[paste0("indoor_", datatype)]
        # There are some columns in indoordata that do not exist in outdoordata
        outdoor_idx <- paste0("outdoor_", datatype)
        if (outdoor_idx %in% names(joineddata)) {
          outdoor <- joineddata[outdoor_idx]
        } else {
          next
        }
        # If indoor and outdoor columns are both of type double, create ratio col
        if (sapply(indoor, class) == "numeric") {
            ratio_name <- paste0("ratio_", datatype)
            joineddata[ratio_name] <- indoor / outdoor
        }
    }
}
#joineddata["ratio_opc_bin0"]
```


```{r}
# Create a mask for before and after filtration

joineddata$hepa_installed <- joineddata$date >= hepatime
#head(joineddata$hepa_installed)
#tail(joineddata$hepa_installed)


#Create new dataframes for HEPA installed and not installed

# hepa_installed <- joineddata[joineddata$date >= hepatime, ]
# no_hepa_installed <- joineddata[joineddata$date < hepatime, ]



#head(no_hepa_installed)
#tail(no_hepa_installed)




```


```{r}
# Find statistics for before and after

# Create summary preparation dataframe, which formats joineddata correctly to
# get summary statistics
summ_prep <- joineddata[, sapply(joineddata, is.numeric)]
summ_prep$hepa_installed <- joineddata$hepa_installed

# Get rid of second underscores
for (colname in names(summ_prep)) {
  # Assume that there can only be 2 underscores max
  chars <- strsplit(colname, "_")[[1]]
  if (length(chars) == 3) {
    # Create new column with simpler name and delete old one
    summ_prep[paste0(chars[1], "_", chars[2], chars[3])] = summ_prep[colname]
    summ_prep[colname] <- NULL
  }
}

# Summarizes data
summary <-
  summ_prep %>%
    group_by(hepa_installed) %>%
    summarize(across(starts_with("indoor"),
                    list(fifth = ~ quantile(.x, .05, na.rm = TRUE),
                          twentyfifth = ~ quantile(.x, .25, na.rm = TRUE),
                          seventyfifth = ~ quantile(.x, .75, na.rm = TRUE),
                          nintyfifth = ~ quantile(.x, .95, na.rm = TRUE),
                          mean = ~ mean(.x, na.rm = TRUE),
                          median = ~ median(.x, na.rm = TRUE)
                          )
                    )
              ) %>%
    pivot_longer(
      names_to = c(NA, "bin", ".value"),
      names_sep = "_",
      cols = -c(hepa_installed)
    )

# For sake of memory, delete summ_prep
summ_prep <- NULL
summary
```


```{r}
# EVERYTHING BELOW IS OLD CODE USED FOR PREVIOUS ANALYSIS
```




Generate truncated dataset for easier analysis.
```{r}
mydata.working <- data.frame(date, ws, wd, pm1, pm25, pm10, co, bin0, bin6, bin11, noise, temp_manifold, rh_manifold)
attach(mydata.working)
names(mydata.working)
header<-names(mydata.working)
```



Plot probability density for quick view.
```{r}

#meltdf<-melt(mydata.working, id="date")
#ggplot(meltdf, aes(x=date, y=value, color=variable, group=variable)) + geom_line()



mydata.working %>% keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_density()   



```
Quick time series for a preliminary view
```{r}
#timePlot(selectByDate(mydata.working, year = 2019, month = "sep"),
#pollutant = c('co', 'bin0', 'bin6' , 'bin11'))

#for(i in 1:ncol(mydata.working)) {
#  TimePlot(mydata.working, pollutant = names(mydata.working[i])
#}

length<-ncol(mydata.working)
pollutants<-numeric(length)
for(i in 2:length) {   #The counter declaration
  pollutants[i]<-names(mydata.working[i])
  #then do something for each iteration
  #timePlot(mydata.working, pollutant=names(mydata.working[i]))
  #....
} #And finish the enviroment.

timePlot(mydata.working, pollutant=pollutants)
#timePlot(mydata.working, pollutant=names(mydata.working[i]))
#lapply(2:ncol(mydata.working),FUN=myPlot)


#timePlot(mydata.working, pollutant = c("co", "bin0"), y.relation = "free")
#timePlot(mydata.working, pollutant = c('co', 'bin0', 'bin6' , 'bin11'), y.relation = "free")

#timePlot(mydata.working, pollutant = "co")
#timePlot(mydata.working, pollutant = "bin0")
#summaryPlot(mydata.working, percentile = 0.99, period="months")
#summaryPlot(subset(mydata, select = c(date, co, bin0, pm1)))
```

Do any cleaning necessary from viewing the quick view - select a date range, eliminate outliers, etc.
```{r}
names(mydata.working)
daterange <- mydata.working$date >"2019-09-01" & mydata.working$date < "2019-10-03"
summary(daterange)
mydata.working.daterange <- mydata.working(daterange)
summaryPlot(mydata.working.daterange)
```


Show correlation plot: create new dataframe that is numeric; ignore NaNs.
```{r}

#mydata.cor <- mydata.working
#mydata.cor$date<-NULL
#summary(mydata.cor)
#working.cor<-cor(na.omit(mydata.cor))
working.num<-mydata.working[, (sapply(mydata.working,is.numeric))]
working.cor<-cor(na.omit(working.num))
corrplot(working.cor)
```

Start plotting specific R figures
```{r}
#polarPlot(mydata.working, pollutant = 'co', col = "jet", key.position = "bottom", key.footer = NULL)
#polarPlot(mydata.working, pollutant = 'bin0', col = "jet", key.position = "bottom", key.footer = NULL)
#polarPlot(mydata.working, pollutant = 'bin6', col = "jet", key.position = "bottom", key.footer = NULL)
#polarPlot(mydata.working, pollutant = 'bin11', col = "jet", key.position = "bottom", key.footer = NULL)



for(i in 1:ncol(mydata.working)){
  polarPlot(mydata.working, pollutant = header[i], col = "jet", key.header = header[i], key.position = "bottom", key.footer = NULL)
  
}
```

Now plot deeper things like k cluster analysis.
```{r}
polarCluster(mydata.working, pollutant="co", n.clusters=2:10, cols= "Set2")
```

Decide on a number of clusters and re-plot that.
```{r}
polarCluster(mydata.working, pollutant = "bin0", x="ws", wd = "wd", n.clusters = 5, cols ="Paired")
```

