# An Algorithm to Detect Decay Constants

This is a demo pipeline that takes a time series of indoor/outdoor data, and
calculate the exponential decay constant of a pollutant after a plume.

When a source releases pollutants into the air and finishes, the
concentration of pollution naturally decays over time. These decay
functions are exponential in nature and can be fitted to the following model:

    A = yf + (y0 - yf)e^(-kt)

Where t is time, y0 is the initial value of the decay, yf is the asymptotic
value of the decay, A is the concentration of PM at time t, and k is the
decay constant.

To test the efficacy of HEPA air purifiers, looking at the decay constants of
these curves before and after a purifier is installed can be important. A
purifier that proves to increase the decay constant, i.e. cause the pollutants
in the air to diffuse out of a space at a faster rate, can be argued to be
effective.

```{r setup}
library(tidyverse)
library(reticulate)
library(pracma)
source("scripts/curve_fitting.R")
# Python functions
py_install("pandas")
source_python("read_pickle.py")
```

Firstly, we will download a random data file for analysis containing indoor and
outdoor air quality data.

```{r}
# Download a file for analysis
df_sensor_sn <- read_pickle_file(
    "hepa-pckls/indoor-outdoor-pairs/302 Lexington Street /MOD-PM-00128.pckl"
)
```

Since this data does not include an indoor/outdoor ratio metric, first that has
to be created. For visual purposes, this code block plots that data as a time
series.

```{r}
# Get indoor/outdoor ratio and add that to dataframe
df_sensor_sn$indoor_opc_pm25 <- as.numeric(df_sensor_sn$indoor_opc_pm25)
df_sensor_sn$outdoor_opc_pm25 <- as.numeric(df_sensor_sn$outdoor_opc_pm25)
df_sensor_sn$in_out_opc_pm25 <-
    df_sensor_sn$indoor_opc_pm25 / df_sensor_sn$outdoor_opc_pm25

# Create separate variable for this time series for ease of use
in_out <- df_sensor_sn$in_out_opc_pm25

# Plot the time series
plot(df_sensor_sn$timestamp, in_out,
  type = "l",
  main = "Indoor/Outdoor Ratio",
  xlab = "Time",
  ylab = "Ratio of indoor/outdoor concentrations",
  col = "navy")
  grid()
```

Next, we must get the peaks and valleys of the data set.

```{r}
# Get peaks of dataset
in_out[is.na(in_out)] <- 0
peaks <- findpeaks(in_out,
                   nups = 1,
                   ndowns = 1,
                   minpeakheight = 10,
                   minpeakdistance = 200,
                   threshold = 0)
print(peaks)

# Plot the peaks on top of current time series
points(peaks[, 2], peaks[, 1], pch = 20, col = "maroon")
```

```{r}
# Get nearest valleys to those peaks
valleys <- get_first_valleys(in_out, peaks, 2)
print(valleys)

# Plot valleys on top of current time series
points(valleys[, 2], valleys[, 1], pch = 20, col = "green")
```

With these peaks and valleys, we now know the edges of our exponential decay
curves. We can then isolate that data and caculate a fit function, or a
"best fit line". The decay constant from that formula can be used for learning
more about the efficacy of the HEPA purifier used here.

```{r}
# Get the decay values between each peak and valley highlighted
decays <- curve_fitting(in_out, peaks, valleys)
decays
```